---
title: "Longitudinal final project"
author: "Emily Saks"
date: "2024-10-15"
output:
  pdf_document: default
  html_document: default
---

Load data and packages and extract the variables of interest
```{r}
# load packages
library(haven)
library(dplyr)
library(ggplot2)
library(tidyr)
library(finalfit)
library(psych)
library(glmmTMB)
library(lme4)
library(car)

# load data. n=935
full_dataset <- read_por("path/to/data.por")
head(full_dataset)

# extract variables
participation_variables<-c("V515", "V1634", "V5850", "V521", "V1642", "V5855", "V526", "V1649", "V5857", "V531", "V1701", "V5859", "V502", "V1621", "V5800")
polarization_variables<-c("V432", "V440", "V1520", "V1525")
knowledge_variables<-c("V760")
ID_variable<-c("V6")
control_variables<-c("V232", "V141", "V76", "V231", "V84", "V767")
variable_names<-c(participation_variables, polarization_variables, knowledge_variables, control_variables, ID_variable)
polpart<- full_dataset[variable_names]
head(polpart)

summary(polpart)
```

Rename, recode and construct variables
```{r}
# rename
polpart<- polpart %>%
  rename("part_persuade_2"= "V515",
         "part_persuade_3" = "V1634",
         "part_persuade_4" = "V5850",
         "part_rallies_2" = "V521",
         "part_rallies_3" = "V1642",
         "part_rallies_4" = "V5855",
         "part_work_2" = "V526",
         "part_work_3" =  "V1649",
         "part_work_4" = "V5857",
         "part_button_2" =  "V531",
         "part_button_3" =  "V1701",
         "part_button_4" =  "V5859",   
         "part_vote_2" = "V502",
         "part_vote_3" = "V1621",
         "part_vote_4"= "V5800")

polpart<-polpart %>%
  rename(
    "ID" = "V6",
    "education"= "V767",
    "partisanship" = "V84",
    "sex" = "V231",
    "parental_soc" = "V76",
    "pol_int" = "V141",
    "race" = "V232"
  )

polpart<-polpart %>%
  rename(
   "pol_know_2" = "V760"
  )

polpart<-polpart %>%
  rename(
    "therm_rep_2" = "V440",
    "therm_dem_2" = "V432",
    "therm_rep_3" = "V1525",
    "therm_dem_3" = "V1520"
  )

# to compute participation index, change to factors and relevel so 1=yes and 0=no
polpart$part_persuade_2<-as.factor(polpart$part_persuade_2)
polpart$part_rallies_2<-as.factor(polpart$part_rallies_2)
polpart$part_work_2<-as.factor(polpart$part_work_2)
polpart$part_button_2<-as.factor(polpart$part_button_2)
polpart$part_vote_2<-as.factor(polpart$part_vote_2)
polpart$part_persuade_3<-as.factor(polpart$part_persuade_3)
polpart$part_rallies_3<-as.factor(polpart$part_rallies_3)
polpart$part_work_3<-as.factor(polpart$part_work_3)
polpart$part_button_3<-as.factor(polpart$part_button_3)
polpart$part_vote_3<-as.factor(polpart$part_vote_3)
polpart$part_persuade_4<-as.factor(polpart$part_persuade_4)
polpart$part_rallies_4<-as.factor(polpart$part_rallies_4)
polpart$part_work_4<-as.factor(polpart$part_work_4)
polpart$part_button_4<-as.factor(polpart$part_button_4)
polpart$part_vote_4<-as.factor(polpart$part_vote_4)
levels(polpart$part_persuade_2)<-c(1,0)
levels(polpart$part_rallies_2)<-c(1,0)
levels(polpart$part_work_2)<-c(1,0)
levels(polpart$part_button_2)<-c(1,0)
levels(polpart$part_vote_2)<-c(1,0, NA)
levels(polpart$part_persuade_3)<-c(1,0)
levels(polpart$part_rallies_3)<-c(1,0)
levels(polpart$part_work_3)<-c(1,0)
levels(polpart$part_button_3)<-c(1,0)
levels(polpart$part_vote_3)<-c(1,0, NA)
levels(polpart$part_persuade_4)<-c(1,0)
levels(polpart$part_rallies_4)<-c(1,0)
levels(polpart$part_work_4)<-c(1,0)
levels(polpart$part_button_4)<-c(1,0)
levels(polpart$part_vote_4)<-c(1,0, NA)
polpart$part_persuade_2<-as.numeric(polpart$part_persuade_2)-1
polpart$part_rallies_2<-as.numeric(polpart$part_rallies_2)-1
polpart$part_work_2<-as.numeric(polpart$part_work_2)-1
polpart$part_button_2<-as.numeric(polpart$part_button_2)-1
polpart$part_vote_2<-as.numeric(polpart$part_vote_2)-1
polpart$part_persuade_3<-as.numeric(polpart$part_persuade_3)-1
polpart$part_rallies_3<-as.numeric(polpart$part_rallies_3)-1
polpart$part_work_3<-as.numeric(polpart$part_work_3)-1
polpart$part_button_3<-as.numeric(polpart$part_button_3)-1
polpart$part_vote_3<-as.numeric(polpart$part_vote_3)-1
polpart$part_persuade_4<-as.numeric(polpart$part_persuade_4)-1
polpart$part_rallies_4<-as.numeric(polpart$part_rallies_4)-1
polpart$part_work_4<-as.numeric(polpart$part_work_4)-1
polpart$part_button_4<-as.numeric(polpart$part_button_4)-1
polpart$part_vote_4<-as.numeric(polpart$part_vote_4)-1

# add participation index
polpart$participation_index_2<-polpart$part_persuade_2+polpart$part_rallies_2+polpart$part_work_2+polpart$part_button_2+polpart$part_vote_2

polpart$participation_index_3<-polpart$part_persuade_3+polpart$part_rallies_3+polpart$part_work_3+polpart$part_button_3+polpart$part_vote_3

polpart$participation_index_4<-polpart$part_persuade_4+polpart$part_rallies_4+polpart$part_work_4+polpart$part_button_4+polpart$part_vote_4

# recode partisanship 
polpart<- polpart %>%
  mutate(ideo= case_when(partisanship== 11 ~ "Democrat",
                         partisanship== 12 ~ "Democrat",
                         partisanship== 13 ~ "Democrat",
                         partisanship== 15 ~ "Republican",
                         partisanship== 16 ~ "Republican",
                         partisanship== 17 ~ "Republican",
                         partisanship== 14 ~ "Neither",
                         partisanship== 20 ~ "Neither",
                         partisanship== 30 ~ "Neither",
                         partisanship== 40 ~ "DK",
                         partisanship== 88 ~ "DK",
                         TRUE ~ "NA"))
polpart$ideo<-as.factor(polpart$ideo)
polpart$ideo<-plyr::revalue(polpart$ideo, c("1"="Democrat", "2"="Neither", "3"="Republican", "99"="NA"))

# recode polint
polpart<- polpart %>%
  mutate(pol_int= case_when(pol_int== 1 ~ "4",
                         pol_int== 2 ~ "3",
                         pol_int== 3 ~ "2",
                         pol_int== 4 ~ "1",
                         TRUE ~ "NA"))


# recode race
polpart<- polpart %>%
  mutate(race= case_when(race== 1 ~ "White",
                         race== 2 ~ "Non-white",
                         race== 3 ~ "Non-white",
                         race== 8 ~ "Non-white",
                         TRUE ~ "NA"))
polpart$race<-as.factor(polpart$race)

# Turn other predictors/controls into factors
to_convert<-c("education", "sex", "partisanship", "pol_int", "pol_know_2")
polpart[, to_convert]<-lapply(polpart[, to_convert], as.factor) # turn into factor

# add in affective polarization measure as difference between inparty and outparty feeling thermometer
polpart<- polpart %>%
  mutate(polarization= case_when(
    partisanship %in% c(11,12,13) ~ abs(therm_dem_2-therm_rep_2),
    partisanship %in% c(15,16,17) ~ abs(therm_rep_2-therm_dem_2),
    TRUE~NA
  )
  )
head(polpart)
sum(is.na(polpart$polarization)) 
```

Examine "DK" and "Neither" responders on outcome variable which is participation index 4
```{r}
# For the participation items, 8=DK and 9=NA.Calculate DK and NAs on participation index 2-4.
sum(is.na(polpart$part_persuade_2)) # 3 NA
sum(is.na(polpart$part_rallies_2)) # 3 NA
sum(is.na(polpart$part_work_2)) # 4 NA
sum(is.na(polpart$part_button_2)) # 2 NA
sum(is.na(polpart$part_vote_2)) 
nrow(subset(polpart, polpart$part_persuade_2==8)) 
nrow(subset(polpart, polpart$part_rallies_2==8))
nrow(subset(polpart, polpart$part_work_2==8))
nrow(subset(polpart, polpart$part_button_2==8))
nrow(subset(polpart, polpart$part_vote_2==8)) # 0 DKs on all

sum(is.na(polpart$part_persuade_3)) # 3 NA
sum(is.na(polpart$part_rallies_3)) # 2 NA
sum(is.na(polpart$part_work_3)) # 2 NA
sum(is.na(polpart$part_button_3)) # 2 NA
sum(is.na(polpart$part_vote_3)) # 2 NA
nrow(subset(polpart, polpart$part_persuade_3==8)) 
nrow(subset(polpart, polpart$part_rallies_3==8))
nrow(subset(polpart, polpart$part_work_3==8))
nrow(subset(polpart, polpart$part_button_3==8))
nrow(subset(polpart, polpart$part_vote_3==8)) # 2 DKs on this last item only

sum(is.na(polpart$part_persuade_4)) # 2 NA
sum(is.na(polpart$part_rallies_4)) # 2 NA
sum(is.na(polpart$part_work_4)) # 2 NA
sum(is.na(polpart$part_button_4)) # 2 NA
sum(is.na(polpart$part_vote_4)) # # 2 NA
nrow(subset(polpart, polpart$part_persuade_4==8))
nrow(subset(polpart, polpart$part_rallies_4==8))
nrow(subset(polpart, polpart$part_work_4==8))
nrow(subset(polpart, polpart$part_button_4==8))
nrow(subset(polpart, polpart$part_vote_4==8)) # no DKs


# Since there are only 2 DK responses and because past literature excludes them, we will exclude the DKs
polpart<-subset(polpart, part_vote_3!=8)


# For missingness, we will examine whether there are any patterns in the missingness to be aware of on the response variable (index at times 2-4)

ff_glimpse(polpart) # shows us percent missing on all variables

polpart %>%
  missing_plot() # examine pattern of missingness
 
explanatory = c("polarization", "education", "parental_soc", "pol_int", "race", "sex", "partisanship", "pol_know_2")
dependent = c("participation_index_4", "participation_index_3", "participation_index_2")

polpart %>%
  missing_pattern(dependent, explanatory) # look at pattern of missingness
polpart %>%
  missing_pairs(dependent, explanatory) # possibly party ID, education, knowledge

# let's confirm the differences. 
explanatory = c("polarization", "education", "parental_soc", "pol_int", "race", "sex", "partisanship", "pol_know_2", "participation_index_2", "participation_index_3")
dependent = c("participation_index_4")
missing_compare(.data=polpart, dependent=dependent, explanatory=explanatory)
polpart %>%
  missing_compare(.data=polpart, dependent=dependent, explanatory=explanatory) %>%
                    knitr::kable(row.names=FALSE)
# No statistically significant differences were detected on the variables examined

explanatory = c("polarization", "education", "parental_soc", "pol_int", "race", "sex", "partisanship", "pol_know_2", "participation_index_2")
dependent = c("participation_index_3")
missing_compare(.data=polpart, dependent=dependent, explanatory=explanatory)
polpart %>%
  missing_compare(.data=polpart, dependent=dependent, explanatory=explanatory) %>%
                    knitr::kable(row.names=FALSE) # we can see no significant p-vales, so we will consider outcome MAR

explanatory = c("polarization", "education", "parental_soc", "pol_int", "race", "sex", "partisanship", "pol_know_2")
dependent = c("participation_index_2")
missing_compare(.data=polpart, dependent=dependent, explanatory=explanatory)
```
\clearpage
```{r}

polpart %>%
  missing_compare(.data=polpart, dependent=dependent, explanatory=explanatory) %>%
                    knitr::kable(row.names=FALSE) 


# We will also examine patterns of missingness on the predictor, affective polarization since there are many (262) of them
explanatory = c("therm_rep_2", "therm_dem_2", "education", "parental_soc", "pol_int", "race", "sex", "partisanship", "pol_know_2", "participation_index_2", "participation_index_3", "participation_index_4")
dependent = c("polarization")
missing_compare(.data=polpart, dependent=dependent, explanatory=explanatory)
polpart %>%
  missing_compare(.data=polpart, dependent=dependent, explanatory=explanatory) %>%
                    knitr::kable(row.names=FALSE) # We can see a significant difference by sex, political knowledge and parental socialization

# final look
summary(polpart)

```

Reshape data
```{r}
# reshaping
polpart.long<-polpart %>%
  pivot_longer(c(1:20, 28:30),
               names_to=c(".value", "Time"),
               names_pattern="(.+)_(.)"
               )
head(polpart.long)

# recode time variable
#  Data was collected 1973, 1982 and 1997. We will code the first interval as 0, the second as 9 and the third as 24
polpart.long<- polpart.long %>%
  mutate(Year= case_when(Time== 2 ~ "0",
                         Time== 3 ~ "9",
                         Time== 4 ~ "24",
                         TRUE ~ "NA"))
polpart.long$Year<-as.numeric(polpart.long$Year)


summary(polpart.long)

# check missing data on the long version of the index
explanatory = c("polarization", "therm_rep", "therm_dem", "education", "parental_soc", "pol_int", "race", "sex", "partisanship", "pol_know")
dependent = c("participation_index")
missing_compare(.data=polpart.long, dependent=dependent, explanatory=explanatory)
polpart %>%
  missing_compare(.data=polpart.long, dependent=dependent, explanatory=explanatory) %>%
                    knitr::kable(row.names=FALSE)
# Education, parental soc, political interest, sex, political knowledge. Acknowledge as a limitation

# could do FIML or multiple imputation

```

Plot the data and look at descriptive statistics
```{r}
# descriptives
describe(polpart.long)

grouptimemeans<-aggregate(x=polpart.long$participation_index,
by=list(polpart.long$Year), FUN=mean,na.rm=TRUE)
colnames(grouptimemeans)<-c("Year", "Political Participation")
knitr::kable(grouptimemeans, digits=3, caption="Means for Political Participation by Wave")
groupsd<-aggregate(x=polpart.long$participation_index,
by=list(polpart.long$Year), FUN=sd,na.rm=TRUE)
colnames(groupsd)<-c("Year", "Political Participation")
knitr::kable(groupsd, digits=3, caption="SD for Political Participation by Wave")

# look at distribution of the outcome variable
hist(polpart$participation_index_2, 
     main="Histogram of Political Participation Wave 2",
     xlab="Political Participation Scores")
hist(polpart$participation_index_3, 
     main="Histogram of Political Participation Wave 3",
     xlab="Political Participation Scores")
hist(polpart$participation_index_4, 
     main="Histogram of Political Participation Wave 4",
     xlab="Political Participation Scores")

# initial plot of individual trajectories
g1 <- ggplot(data = polpart.long, aes(x=Year,y=participation_index, group=ID))+
  geom_line()+
  theme_bw ()+
  scale_y_continuous(name="Political Participation")+
  scale_linetype(name="Polarization")+
  ggtitle("Individual Trajectories for Political Participation Over Time")
g1


# draw on fitted linear regression lines
g2 <- ggplot(data = polpart.long,aes(x=Year,y=participation_index))+ 
  geom_point () + 
  theme_bw () +
  stat_summary(fun="mean",geom="point",size=6,shape=18,colour="blue")+
  stat_summary(fun="mean",geom="line",lwd=1.2,colour="blue")+
  stat_smooth(method="lm",se=FALSE,lwd=1.2,formula=y~x,colour="red")+   scale_y_continuous(name="Political Participation")+
  ggtitle("Mean Trajectory of Political Participation Over Time")
g2

# examine by polarization
polpart.long$polarization_cut<-cut_interval(polpart.long$polarization, n=4)
g3<-ggplot(data=polpart.long, aes(x=Year, y=participation_index, group=ID))+
  geom_line()+
  stat_summary(fun=mean,geom="line",lwd=3,aes(group=1))+
  facet_grid(.~polarization_cut)
g3

```

Begin testing baseline model -- Just the time variable (but can be linear, quadratic, etc.) predictor, and outcome variable. Use model comparison to see whether you need both time and predictor and which ones have a random component in addition to a fixed component. 
```{r}
## Fit different baseline models ##

# Begin with testing negative binomial since it is a count index
# we can observe that the mean of our outcome variable 3.07 is much lower than its variance 19
#Assumptions: independence, homoskedasticity is no longer assumed because variance is a function of the mean, the model is linear in parameters but expected response is not linearly related to them
#baseline_nb1<-glmer.nb(participation_index ~ 1 + Year + (1|ID), data=polpart.long)
#summary(baseline_nb1) 
baseline_nb2<-glmmTMB(participation_index ~ 1 + Year + (1 + Year|ID), data=polpart.long, family=nbinom2, control=glmmTMBControl(optimizer=optim, optArgs=list(method="BFGS")))
summary(baseline_nb2) # needed to change optimizer due to convergence issues

# Poisson
baseline_poisson<- glmmTMB(participation_index ~ 1 + Year + (1 + Year|ID), data=polpart.long, family=poisson, control=glmmTMBControl(optimizer=optim, optArgs=list(method="BFGS"))) 
summary(baseline_poisson)

# Linear 
baseline_linear1<-lmer(participation_index~ 1 + Year + (1 + Year|ID),
             data=polpart.long,
             control = lmerControl(optimizer ="Nelder_Mead",
                                  check.conv.singular = .makeCC(action = "ignore",  tol = 1e-3)),
             REML=FALSE)
summary(baseline_linear1) 


# Compare to model with only random slope
baseline_linear2<-lmer(participation_index~ 1 + Year + (0 + Year|ID),
             data=polpart.long,
             REML=FALSE)
summary(baseline_linear2) # AIC=17418.5, BIC=17442.2 

# Compare to model with only random intercept
baseline_linear3<-lmer(participation_index~ 1 + Year + (1|ID),
             data=polpart.long, control = lmerControl(optimizer ="Nelder_Mead",
                                  check.conv.singular = .makeCC(action = "ignore",  tol = 1e-3)),
             REML=FALSE)
summary(baseline_linear3) # needed to change optimizer

# Compare to model with only fixed effects
baseline_linear4<-lm(participation_index~ 1 + Year,
             data=polpart.long)
summary(baseline_linear4) 
BIC(baseline_linear4)



```
Figure comparing models
```{r}
## Figure comparing functional forms##
# fixed intercept, fixed slope, variance of random intercept, variance of random slope, error variance, BIC
modnbinom<-c("1.259", "-0.013", "6.995e-03", "2.176e-05 ",  "9311.800")
modpos<-c("1.259", "-0.013", "6.891e-03", "2.166e-05", "9303.900")
modlin1<-c("3.567", "-0.039", "0.577", "0.000",  "8452.200")

table1<-cbind(modnbinom, modpos, modlin1)
rownames(table1)<-c("Fixed Intercept", "Fixed Slope", "Variance of Random Intercept", "Variance of Random Slope", "BIC")
knitr::kable(table1, digits=3, caption="Parameter Estimates for Functional Forms", col.names=c("Negative Binomial", "Poisson", "Linear"), align="c")


## Figure comparing fixed vs. random effects##
modlin2<-c("3.507", "-0.037", "-", "0.002", "1.342", "8715.500")
modlin3<-c("3.581", "-0.040", "0.823", "-", "1.046", "8471.300")
modlin4<-c("3.507", "0.037", "-", "-", "1.860", "8870.164")
table2<-cbind(modlin1, modlin2, modlin3, modlin4)
rownames(table2)<-c("Fixed Intercept", "Fixed Slope", "Variance of Random Intercept", "Variance of Random Slope", "Error Variance", "BIC")
knitr::kable(table2, digits=3, caption="Parameter Estimates for Determining Inclusion of Random Effects", col.names=c("Random Intercept and Random Slope", "Fixed Intercept and Random Slope", "Random Intercept and Fixed Slope", "Fixed Intercept and Fixed Slope"), align="c")
```


Check assumptions
```{r}
# Are random effects normally distributed?
#fixef(baseline_linear1)
#ranef(baseline_linear1)
ind.intercepts<-fixef(baseline_linear1)[1] + ranef(baseline_linear1)[[1]][, 1]
hist(ind.intercepts)
boxplot(ind.intercepts)
qqnorm(ind.intercepts)
qqline(ind.intercepts)
describe(ind.intercepts) # yes, look normal

ind.slopes<- fixef(baseline_linear1)[2] + ranef(baseline_linear1)[[1]][,2] # get individual slopes
hist(ind.slopes)
boxplot(ind.slopes)
qqnorm(ind.slopes)
qqline(ind.slopes)
describe(ind.slopes)

# Are the residuals/errors normally distributed with a mean of 0?
res<-resid(baseline_linear1)
boxplot(res)
qqnorm(res)
qqline(res)
describe(res)

# Are the errors independent?
plot(x=polpart.long$Year, y=res[1:length(polpart.long$Year)], xlab="Time", ylab="Residual")

``` 

Specify the variance-covariance matrix for repeated measures residuals and the random effects error structure.
```{r}
# Check to see if we can assume an independent error structure ##

# compute covariance matrix
polpart<-as.data.frame(polpart)
table5<-round(cov(x=polpart[, 28:30], use="pairwise.complete.obs"), digits=3)
rownames(table5)<-c("Time 1", "Time 2", "Time 3")
knitr::kable(table5, digits=3, caption="Covariance Matrix", colnames=c("Time 1", "Time 2", "Time 3"))

# compute correlation matrix of residuals at each time point
polpart.long2<-subset(polpart.long, is.na(polpart.long$participation_index)==FALSE)
polpart.long2$fitted_mod<-na.omit(fitted(baseline_linear1))
polpart.long2$resid_mod<-resid(baseline_linear1)
resid_0<-polpart.long2[polpart.long2$Year==0, ]$resid_mod
resid_1<-polpart.long2[polpart.long2$Year==9, ]$resid_mod
resid_2<-polpart.long2[polpart.long2$Year==24, ]$resid_mod
residuals<- cbind(resid_0, resid_1, resid_2)
round(residuals,3)
table6<-round(cor(residuals), digits=3)
rownames(table6)<-c("Time 1", "Time 2", "Time 3")
knitr::kable(table6, digits=3, caption="Correlation Matrix of Residuals", col.names=c("Time 1", "Time 2", "Time 3"))

# We can see that the residuals are not highly correlated between time points. Also, the variance of scores between time points is fairly similar

# We could test alternative error structures but I think it meets assumptions for independent

```

Test covariates
```{r}

# Baseline model
baseline_linear1<-lmer(participation_index~ 1 + Year + (1 + Year|ID),
             data=polpart.long,
             control = lmerControl(optimizer ="Nelder_Mead",
                                  check.conv.singular = .makeCC(action = "ignore",  tol = 1e-3)),
             REML=FALSE)
summary(baseline_linear1) 

# Model with demographic covariates
linear1<-lmer(participation_index~ 1 + Year + race + sex + education + partisanship + (1 + Year|ID),
             data=polpart.long,
             control = lmerControl(optimizer ="Nelder_Mead",
                                  check.conv.singular = .makeCC(action = "ignore",  tol = 1e-3)),
             REML=FALSE)
summary(linear1)

# Model with pol int and parental soc
linear2<-lmer(participation_index~ 1 + Year + race + sex + education + partisanship + pol_int + parental_soc + (1 + Year|ID),
             data=polpart.long,
             control = lmerControl(optimizer ="Nelder_Mead",
                                  check.conv.singular = .makeCC(action = "ignore",  tol = 1e-3)),
             REML=FALSE)
summary(linear2)

# multimodel inference
library("bbmle")                              

mynames <- as.character(1:3) 								# model numbers as characters
myaicc  <- ICtab(logLik(baseline_linear1), 
		             logLik(linear1), 
		             logLik(linear2), 
		             type ="AICc",
		             nobs = nrow(na.omit(polpart.long)),
		             weights =TRUE,delta =TRUE,
		             sort=TRUE,mnames=mynames,base=TRUE)

myaicc$rweight <- max(myaicc$weight)/myaicc$weight
myaicc

# we can see that adding in the controls improves the model
```


Test hypotheses
```{r}
# Model with hypothesized independent variable polarization
linear3<-lmer(participation_index~ 1 + Year + race + sex + education + partisanship + pol_int + parental_soc + polarization + (1 + Year|ID),
             data=polpart.long,
             control = lmerControl(optimizer ="Nelder_Mead",
                                  check.conv.singular = .makeCC(action = "ignore",  tol = 1e-3)),
             REML=FALSE)
summary(linear3)

# model with interaction
linear4<-lmer(participation_index~ 1 + Year*polarization + race + sex + education + partisanship + pol_int + parental_soc  + (1 + Year|ID),
             data=polpart.long,
             control = lmerControl(optimizer ="Nelder_Mead",
                                  check.conv.singular = .makeCC(action = "ignore",  tol = 1e-3)),
             REML=FALSE)
summary(linear4)


# multimodel inference
mynames <- as.character(1:5) 							
myaicc  <- ICtab(logLik(baseline_linear1), 
		             logLik(linear1), 
		             logLik(linear2), 
		             logLik(linear3), 
		             logLik(linear4),
		             type ="AICc",
		             nobs = nrow(na.omit(polpart.long)),
		             weights =TRUE,delta =TRUE,
		             sort=TRUE,mnames=mynames,base=TRUE)

myaicc$rweight <- max(myaicc$weight)/myaicc$weight
myaicc

# Calculate R^2
library(MuMIn)
r.squaredGLMM(linear3)
r.squaredGLMM(linear4)
# Marginal only takes into account fixed effects. The marginal R^2 is 0.17 and the conditional is 0.47, which is good in psychology.
```
Figures with final model comparison
```{r}
modelsummary::msummary(list("Model 1"=linear2, "Model 2"=linear3, "Model 3"=linear4), output="flextable", coef_omit=c(19:22))
```

